---
title: "demuxlet results: comparison with previous samples"
author: "Lukas Weber"
date: "2/18/2021"
output:
  html_document:
    toc: true
    toc_depth: 3
---

# demuxlet results

Demultiplexing results from demuxlet for previous single-cell samples (16030X2-X4). For comparison with demuxlet results from new samples.


# Load outputs


## Load demuxlet outputs

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggExtra))
suppressPackageStartupMessages(library(grid))
```


```{r, message=FALSE}
dir_out <- "../outputs/demuxlet_previous_samples/HGSOC/20pc/bulkBcftools_demuxlet"

# demuxlet outputs
out_demuxlet_best <- read_tsv(file.path(dir_out, "demuxlet.best"))
# not needed
#out_demuxlet_single <- read_tsv(file.path(dir_out, "demuxlet.single"))
#out_demuxlet_sing2 <- read_tsv(file.path(dir_out, "demuxlet.sing2"))

# clean group names and combine doublets into a single group
out_demuxlet_best$best_clean <- gsub("AMB-.*", "AMB", 
                                     gsub("DBL-.*", "DBL", out_demuxlet_best$BEST))
```



# Plots

## Summary

### Number of cells per donor

```{r, message=FALSE, fig.width=4.5, fig.height=3.5}
# plot number of cells per donor from demuxlet
df <- as.data.frame(table(out_demuxlet_best$best_clean))

p_out <- ggplot(df, aes(x = Var1, y = Freq, fill = Var1, label = Freq)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(y = max(Freq) / 20)) + 
  labs(x = element_blank(), 
       y = "count", 
       fill = "donor ID", 
       title = paste0("demuxlet"), 
       subtitle = paste0("computational pool: 16030X2-X4, 20pc doublets")) + 
  theme_bw()

print(p_out)

fn <- file.path("..", "plots", "demuxlet_summary_previous_samples.png")
png(fn, width = 5, height = 4, units = "in", res = 200)
print(p_out)
dev.off()
```



## Number of reads overlapping with variants

### Number of reads overlapping with variants per droplet (raw, log10)

```{r, message=FALSE, fig.width=4.5, fig.height=3.75}
# plot number of variants per droplet from demuxlet
p_out <- ggplot(out_demuxlet_best, aes(x = log10(RD.TOTL))) + 
  geom_histogram(fill = "blue3") + 
  labs(title = paste0("demuxlet"), 
       subtitle = paste0("reads overlapping variants: raw\n", 
                         "computational pool: 16030X2-X4, 20pc doublets")) + 
  theme_bw()

print(p_out)

fn <- file.path("..", "plots", "demuxlet_RD.TOTL_previous_samples.png")
ggsave(fn, width = 5, height = 4)
```


### Number of reads overlapping with variants per droplet (reads passing QC, log10)

```{r, message=FALSE, fig.width=4.5, fig.height=3.75}
# plot number of variants per droplet from demuxlet
p_out <- ggplot(out_demuxlet_best, aes(x = log10(RD.PASS))) + 
  geom_histogram(fill = "blue4") + 
  labs(title = paste0("demuxlet"), 
       subtitle = paste0("reads overlapping variants: reads passing QC\n", 
                         "computational pool: 16030X2-X4, 20pc doublets")) + 
  theme_bw()

print(p_out)

fn <- file.path("..", "plots", "demuxlet_RD.PASS_previous_samples.png")
ggsave(fn, width = 5, height = 4)
```


### Number of UMIs overlapping with variants per droplet (UMIs passing QC, log10)

```{r, message=FALSE, fig.width=4.5, fig.height=3.75}
# plot number of variants per droplet from demuxlet
p_out <- ggplot(out_demuxlet_best, aes(x = log10(RD.UNIQ))) + 
  geom_histogram(fill = "navy") + 
  labs(title = paste0("demuxlet"), 
       subtitle = paste0("UMIs overlapping variants: UMIs passing QC\n", 
                         "computational pool: 16030X2-X4, 20pc doublets")) + 
  theme_bw()

print(p_out)

fn <- file.path("..", "plots", "demuxlet_RD.UNIQ_previous_samples.png")
ggsave(fn, width = 5, height = 4)
```


### Number of variants overlapping with any read (log10)

```{r, message=FALSE, fig.width=4.5, fig.height=3.75}
# plot number of variants per droplet from demuxlet
p_out <- ggplot(out_demuxlet_best, aes(x = log10(N.SNP))) + 
  geom_histogram(fill = "darkgreen") + 
  labs(title = paste0("demuxlet"), 
       subtitle = paste0("Number of variants overlapping with reads\n", 
                         "computational pool: 16030X2-X4, 20pc doublets")) + 
  theme_bw()

print(p_out)

fn <- file.path("..", "plots", "demuxlet_N.SNP_previous_samples.png")
ggsave(fn, width = 5, height = 4)
```


